---
title: "R-primer-2"
author: "Akhilesh Tayade"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The notebook doesn't essentially give full-proof implementation of different R utilities as the notebook was uploaded to the Github repository after the concepts were discussed in the lecture.

Rather, one would encounter errors while running the code. Can you think how to solve them? :)

**Function**

``` r
function_name <- function (# needs an argument?) {
  # do something
  # return something
}
```

```{r}

# What does a scope of a variable mean? 
# What's the difference between a local and a global variable?

x <- 2
  
# Function definition
replacef <- function(x) {
  x <- 5
  return(x)
}

x <- replacef(10)
x

# Function call
multiply(c(1, 2, 3))

# Function call 
multiply(rep(2, 50))

func <- multiply
func(0)
```

```{r danger}
x <- 2

foo <- function(x) {
  print(x)
}

foo(4)
```

```{r}
divide <- function(x = 6, y) {
  print(x)
  return(x/y)
}

divide(3)
```

**Conditional expression**

```{r}

x <- "2"

# expression #1
if (x) {
  pri
}

x <- NULL

# expression #2
if (length(x)) {
  print(x)
}

x <- 2

# expression #3
if (x == "I") {
  print(x)
} else {
  print('Not I')
}
```

**Loops**

```{r}
# for loop
x <- 1:2

# Continue iterating
for (i in x) {
  print("It's really hot today!")
  if (i == 1) {
    next
  }
  print("May be it's not!")
}

# How do you break the loop without reaching the end of the iteration?
```

```{r}
# while loop
i <- 0

while (i < 3) {
  print(i)
  i = i + 1
}
```

At the end of this lecture, we saw many examples of how to use `ggplot` from `ggplot2`. Please refer to [Chapter 3, Data Visualisation of "R for data science](https://r4ds.had.co.nz)" to understand how to plot beautiful plots using `ggplot` .
