---
title: "Working with Packages"
output: html_notebook
---

### This notebook aims to explain installing and working with tidyverse packages with some example features implemented on data frames. Have fun!

### Installing Tidyverse package
#### *Tidyverse package contains a list of packages useful for working with data

```{r}
install.packages("tidyverse")
```
### Loading the tidyverse package

```{r}
library(tidyverse)
```
### Using Tibble and Readr packages
#### Let's first load these two packages 

```{r}
library(tibble)
library(readr)
```

#### As an example we are using statewise covid testing data for India
#### Using base R functions

```{r}
# Importing covid testing data 
covid_testdata <- read.csv('StatewiseTestingDetails.csv')
```
```{r}
# Using head function to view first few rows of the data
head(covid_testdata)
```
```{r}
# Using nrow()  and ncol() base R functions to view the number of rows and columns for the data
nrow(covid_testdata)
ncol(covid_testdata)
```
### Now let's try this using tidyverse functions
### We are using tibble package from tidyverse, tibble is a modern package to work with dataframes in a better way than base R functions
```{r}
# converting the data to tibble
covid_data <- as_tibble(covid_testdata)
covid_data
```
#### We saw here that using _tibble_ package made things easier, using base R function we had to separately use _head()_, _nrow()_ and _ncol()_ functions to get the necessary information from our data frame, but these features are by default shown by _tibble_ 

### Now we will look at another tidyverse package - __dplyr__
#### dplyr package is used for editing tibbles 

```{r}
#loading dplyr package
library(dplyr)
```
#### We will look at filter() and select() functions from the dplyr package
#### filter() function works on rows for example if we want to reduce our data only to the days when there were more than 5000 covid positive cases
```{r}
filter(covid_data, Positive > 5000)
```
#### Let's try filter to view data only for Uttarakhand state
```{r}
filter(covid_data, State == 'Uttarakhand')
```
#### Select function is used to select columns 
#### as an example, if we just want to view the states and the total Samples collected 
```{r}
select(covid_data, State, TotalSamples)
```
#### or if we want to view all columns other than the column with number of Negative results, we can use '-' minus to eliminate that column
```{r}
select(covid_data, -Negative)
```
#### Another function mutate an be used to edit the content of the tibble
#### Let's say if we want the proportion of positive cases, we can use mutate as shown below
```{r}
newdata = mutate(covid_data, prop = Positive/TotalSamples)
newdata
```
#### To view it as percentage we can just multiply the same column 'prop' by 100
```{r}
newdata = mutate(newdata, prop = prop*100)
newdata
```
#### arrange function from dplyr package is used to arrange tibble data in an order
#### Let's arrange rows in the ascending order of Date (Note that ascending is the by default order in arrange function)
```{r}
arrange(covid_data, Date)
```
#### Let's arrange rows in the descending order of number of positive cases
```{r}
arrange(covid_data, desc(Positive))
```

```{r}
write.csv(covid_data,file= "ggplot/harsh.csv")
```

