---
title: "Working with Packages -- Tidyverse"
output: html_notebook
---

#### R is a community project with a massive amount of content made available by its active user base. New functions are assembled into libraries called ‘packages’, which can be installed using the install.packages() function.

### This notebook aims to explain installing and working with tidyverse packages with some example features implemented on data frames. Have fun!

### Installing Tidyverse package
#### *Tidyverse package contains a list of packages useful for working with data

```{r}
install.packages("tidyverse")
```
### Loading the tidyverse package

```{r}
library(tidyverse)
```
### Using Tibble and Readr packages:  this is a modern take on data frames. 
#### Let's first load these two packages 

```{r}
library(tibble)
library(readr)
```

#### As an example we are using statewise covid testing data for India
#### Using base R functions

```{r}
# Importing covid testing data 
covid_testdata <- read.csv('StatewiseTestingDetails.csv')
```
```{r}
# Using head function to view first few rows of the data
head(covid_testdata)
```
```{r}
# Using nrow()  and ncol() base R functions to view the number of rows and columns for the data
nrow(covid_testdata)
ncol(covid_testdata)
```
### Now let's try this using tidyverse functions
### We are using tibble package from tidyverse, tibble is a modern package to work with dataframes in a better way than base R functions
```{r}
# converting the data to tibble
covid_data <- as_tibble(covid_testdata)
covid_data
```
#### We saw here that using _tibble_ package made things easier, using base R function we had to separately use _head()_, _nrow()_ and _ncol()_ functions to get the necessary information from our data frame, but these features are by default shown by _tibble_ 

#### so next you might ask how could I directly read a file as a tibble? Easy: all you have to do is use read_csv (instead of read.csv)

```{r}
covid_data <- read_csv('StatewiseTestingDetails.csv')
covid_data
```



### Now we will look at another tidyverse package - __dplyr__
#### dplyr package is used for editing tibbles 

```{r}
#loading dplyr package
library(dplyr)
```
#### We will look at filter() and select() functions from the dplyr package
### Filter
#### filter() function works on rows for example if we want to reduce our data only to the days when there were more than 5000 covid positive cases
```{r}
filter(covid_data, Positive > 5000)
```
#### Let's try filter to view data only for Uttarakhand state
```{r}
filter(covid_data, State == 'Uttarakhand')
```
### Select
#### Select function is used to select columns 
#### as an example, if we just want to view the states and the total Samples collected 
```{r}
select(covid_data, State, TotalSamples)
```
#### or if we want to view all columns other than the column with number of Negative results, we can use '-' minus to eliminate that column
```{r}
select(covid_data, -Negative)
```

#### Or you can select a few contiguous coloumns 

```{r}
select(covid_data, State:Positive)
```


### Rename 
#### Rename can be used to easily rename columns
```{r}
renamed_covid_data <- rename(covid_data, Neg = Negative)
renamed_covid_data
```



### Mutate
#### Another function mutate an be used to edit the content of the tibble
#### Let's say if we want the proportion of positive cases, we can use mutate as shown below
```{r}
newdata = mutate(covid_data, prop = Positive/TotalSamples)
newdata
```
#### To view it as percentage we can just multiply the same column 'prop' by 100
```{r}
newdata = mutate(newdata, prop = prop*100)
newdata
```
### Arrange
#### arrange function from dplyr package is used to arrange tibble data in an order
#### Let's arrange rows in the ascending order of Date (Note that ascending is the by default order in arrange function)
```{r}
arrange(covid_data, Date)
```
#### Let's arrange rows in the descending order of number of positive cases
```{r}
arrange(covid_data, desc(Positive))
```


### Writing, reading files

#### Now you may want to save this modified file in your workspace for future use. 
To do this you can simply use the write command
```{r}
write.csv(covid_data,file= "/cloud/project/covid_data_copy.csv")
```

#### you may then later want to read the file into your workspace. For this you can use the read command.
```{r}
covid_data <- read.csv(file= "/cloud/project/covid_data_copy.csv")
```

#### to find your current directory and see the files in the directory
```{r}
getwd() #current directory
list.files() # list all files in the directory
```


